---
title: "Problem Set 2"
author: "Jason Yoo"
date: "2/12/2020"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading the necessary data}
library(tidyverse) 
library(fivethirtyeight)
library(ggplot2)
library(ggthemes)
library(gov.1005.data)
data("endorsements_2020")
data("congress")
```

1) Earliest date for an endorsement in the dataset is `r endorsements_2020 %>% arrange(date) %>% slice(1) %>% pull(date) `.

2) The most recent date for an endorsement is `r endorsements_2020 %>% arrange(desc(date)) %>% slice(1) %>% pull(date) `.

3) `r endorsements_2020 %>% filter(position=="mayor") %>% filter(endorsee=="Pete Buttigieg") %>% slice(1) %>% pull(endorser) ` is the mayor that has endorsed Pete Buttigieg.

4) Endorsers from New Hampshire have an average of `r endorsements_2020 %>% filter(state=="NH") %>% summarize(avgpoints = mean(points)) %>% slice(1) %>% pull(avgpoints) ` endorsement points. (Learn how fivethirtyeight calculates endorsement points here.)

5) Between them, the two female candidates who remain in the race (Amy Klobuchar and Elizabeth Warren) have a total of `r endorsements_2020 %>% filter(endorsee=="Amy Klobuchar" | endorsee== "Elizabeth Warren") %>% summarize(total=sum(points)) %>% slice(1) %>% pull(total) ` endorsement points.

6) The type of data you are working with (words, numbers, categories, or fancier data types like dates) determines what kinds of operations you can perform on it. To distinguish different types of data, it’s useful to use the function class(). This function allows you to get to know what kind of data you’re working with.

6a. The variable “position” is a `r class(endorsements_2020$position)` class of variable.

6b. The variable “endorser” is a `r class(endorsements_2020$endorser) ` class of variable.

6c. The variable “points” is a `r class(endorsements_2020$points) `numeric class of variable.

```{r creating labels in advance for the axis tick texts}
xlabels <- c("1950", "'60", "'70", "'80", "'90", "2000", "'10")
ylabels <- c("40", "45", "50", "55", "60 yrs")
```

```{r readying the data for plotting in the next code chunk}
age.data <- congress %>% 
  filter(party %in% c("D","R")) %>% 
  group_by(year, party) %>% 
  summarize(avg.age=mean(age))
```

```{r the plot itself!}
ggplot(age.data, aes(year, avg.age, color=party)) + 
  geom_line(show.legend = FALSE) + 
  labs(title="Average Age of Members of Congress", 
      subtitle="At start of term, 1947-2013", 
      caption="source: fivethirtyeight package") + 
  scale_color_manual(values=c('blue','red')) + scale_x_continuous(breaks=seq(1950,2010, 10), 
      limits=c(1947, 2013), labels=xlabels) +
  scale_y_continuous(breaks=seq(40,60, 5), limits=c(40,63), labels=ylabels) +
  annotate("text", x = 1965.7, y = 57, label = "Republicans", color="red") +
  annotate("text", x = 1996.5, y = 58, label = "Democrats", color="blue") +
  theme_fivethirtyeight()
```

