---
title: "Problem Set 2"
author: "Jason Yoo"
date: "2/12/2020"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = F)
```

```{r packages}
# In this code chunk, I load the necessary library packages. 

library(tidyverse) 
library(fivethirtyeight)
library(ggplot2)
library(ggthemes)
library(gov.1005.data)
```

```{r data}
# Then, I imported the datasets, "endorsements_2020" and "congress," so that I can answer the following mad libs.
data("endorsements_2020")
data("congress")
```

```{r question1}
# Here, I used the arrange function and sliced the top row, pulling the date column value out for the answer, q1.

q1 <- endorsements_2020 %>% 
  arrange(date) %>% 
  slice(1) %>% 
  pull(date)
```

1) Earliest date for an endorsement in the dataset is `r q1`.

```{r question2}
# Here, I used the arrange function again, but arranged the columns in descending order according to date.
# Again, I sliced the top row and pull out the date value.

q2 <- endorsements_2020 %>% 
  arrange(desc(date)) %>% 
  slice(1) %>% 
  pull(date)
```

2) The most recent date for an endorsement is `r q2`.

```{r question3}
# Here, I filtered by only the observations with position being mayor.
# Then, I filtered again by mayors who endorsed Pete Buttigieg. 

q3 <- endorsements_2020 %>% filter(position=="mayor") %>% 
  filter(endorsee=="Pete Buttigieg") %>% 
  slice(1) %>% 
  pull(endorser)
```

3) `r q3` is the mayor that has endorsed Pete Buttigieg.

```{r question4}
# Similarly, I filtered by only New Hampshire.
# Next, I summarized by the mean of the points and sliced the average in the first row.

q4 <- endorsements_2020 %>% 
  filter(state=="NH") %>% 
  summarize(avgpoints = mean(points)) %>% 
  slice(1) %>% 
  pull(avgpoints)
```

4) Endorsers from New Hampshire have an average of `r q4` endorsement points. (Learn how fivethirtyeight calculates endorsement points here.)

```{r question5}
# This time, I used the logical "or" statement to filter by those who endorsed Amy Klobuchar or Elizabeth Warren.
# I sliced the first row and pulled the total.

q5 <- endorsements_2020 %>% 
  filter(endorsee=="Amy Klobuchar" | endorsee== "Elizabeth Warren") %>% 
  summarize(total=sum(points)) %>% 
  slice(1) %>% 
  pull(total)
```

5) Between them, the two female candidates who remain in the race (Amy Klobuchar and Elizabeth Warren) have a total of `r q5` endorsement points.

```{r question6}
# Using the "$" sign, I selected the appropriate columns to observe the variable class of each column.

q6a <- class(endorsements_2020$position)
q6b <- class(endorsements_2020$endorser)
q6c <- class(endorsements_2020$points)
```

6) The type of data you are working with (words, numbers, categories, or fancier data types like dates) determines what kinds of operations you can perform on it. To distinguish different types of data, it’s useful to use the function class(). This function allows you to get to know what kind of data you’re working with.

6a. The variable “position” is a `r q6a` class of variable. #can also use class(endorsements_2020 %>% pull(position))

6b. The variable “endorser” is a `r q6b` class of variable.

6c. The variable “points” is a `r q6c` numeric class of variable.

```{r labels}
# For the axis ticks in the graph, I noticed they were custom set in the example. 
# I set custom tick marks in this code chunk.

xlabels <- c("1950", "'60", "'70", "'80", "'90", "2000", "'10")
ylabels <- c("40", "45", "50", "55", "60 yrs")
```

```{r dataready}
# Preparing the data for plotting, filtering by only Democratic and Republican parties and grouping by year and party.
# Creating a variable called avg.age to condense average ages in each year for each party.

age.data <- congress %>% 
  filter(party %in% c("D","R")) %>% 
  group_by(year, party) %>% 
  summarize(avg.age=mean(age))
```

```{r plot}
# Using ggplot2 package, set a plot with year on the x-axis and avg.age on the y-axis, coloring by party.
ggplot(age.data, aes(year, avg.age, color=party)) + 
  
  # Set the plot as a line graph, with a line for each party.
  geom_line(show.legend = FALSE) + 
  
  # Setting the title, subtitle, and source caption on the bottom right of plot.
  labs(title="Average Age of Members of Congress", 
      subtitle="At start of term, 1947-2013", 
      caption="source: fivethirtyeight package") + 
  
  # Switching colors of the parties to correct colors.
  scale_color_manual(values=c('blue','red')) +
  
  # Setting the limits on the x-axis and breaks of 10 years.
  scale_x_continuous(breaks=seq(1950, 2010, 10), 
      limits=c(1947, 2013), labels=xlabels) +
  
  # Setting the limits on the y-axis and breaks of 5 years.
  scale_y_continuous(breaks=seq(40, 60, 5), limits=c(40, 63), labels=ylabels) +
  
  # Creating texts to label each line on the plot.
  annotate("text", x = 1965.7, y = 57, label = "Republicans", color="red") +
  annotate("text", x = 1996.5, y = 58, label = "Democrats", color="blue") +
  
  # Cleaning up the aesthetics to resemble the plot in the example.
  theme_fivethirtyeight()
```

